### 2021-12-16 tex/paper.wecs/corrections_Aluisio/mudancas_texto_DEC2021.tex
Aluisio's corrections

PAGE 1 COLUMN 1

* LINE 24 "to images available at larger and larger from time to time" ==> "to large images available over long periods".

PAGE 1 COLUMN 2

* LINE 7 "not affected" ==> "unaffected"

* LINE 18 "is" XXXXX

* LINE 20 "In more details," XXXXX

* LINES 21-22 "not only ... problems" XXXXXX

* LINES 33-34 "In the light ... concepts" ==> "Given the aforementioned discussion and motivation"

* LINE -2 "A basic" ==> "basic"

PAGE 2 COLUMN 1

PARAGRAPH 3

* LINE 8 'needed" ==> "desirable".

PARAGRAPH 4

* LINE 1 "In details," XXXXX

* LINE 2 "for ultra-high" ==> "for high and ultra-high"

* LINE 7 ", where " XXXXX

* LINE 9 "is that" ==> "is such that"

* LINE 9 "what" ==> "which"

* LINE 10 "unfeasible, but only" ==> "unfeasible. Moreover, only"

* LINE -3 "could" ==> "should"

PAGE 2 COLUMN 2

PARAGRAPH 2

* LINE 9 "along" ==> "over"

PARAGRAPH 3

* LINE 1 "After" ==> "The"

* LINE 2 ", one of its byproducts is" ==> "results in"

* LINE 3 ", of so" ==> "of the so"

PARAGRAPH 4

* LINE 1 "can be" ==> "may be"

* LINE 2 "analysis" ==> "analyses"

* LINES 2-3 "different types of wavelet transform and basis" ==> "different wavelet bases"

* LINE 3 "as well the use of" ==> "as well as the"

* LINE 4 "for detail" ==> "of detail"

PAGE 3 COLUMN 1

PARAGRAPH 2

* LINE 1 "The most right-hand term in " ==> "The last term on the right-hand side of "

* LINE 3 "regarding" ==> "and"

* LINE 5 "the time" ==> "time"

* LINE 8 "the spatial" ==> "spatial"

* LINES 9-10 "the Pearson correlation coefficient rises as an convenient alternative" ==> 
"the Pearson correlation coefficient is an interesting alternative".

* LINE -0 ", as explained earlier" ==> "[26]"

* FORMULAS (2) e (3) Faltam pontos finais.

PARAGRAPH 6

* LINES 1-2  "Consequen ... R ... assigned" ==> "Let R=[R_{kl}], where R_{kl} is assigned"

PARAGRAPH 7

* LINE 1 "Let us define" ==> "Define"

* LINE -2 "that for " ==> "that, for

* ACRESCENTAR AQUI A IDEIA DE REGRESSAO ... ******Está no final do arquivo

PAGE 3 COLUMN 2

PARAGRAPH 2

* LINE 2 "distinct studies" ==> "studies"

PARAGRAPH 3

* LINE 2 "identify" ==> "identifying"

PARAGRAPH 3

* LINES -1-(-2) "Posterior ... out" XXXXX

PARAGRAPH 4

* LINE 1 "Lastly, the" ==> "The"

 PAGE 4 COLUMN 1

 PARAGRAPH 1

* LINES 1-2 "All the changes made that occur among subsequent" ==> "All changes between subsequent"

* LINE 3 "the "ones")" ==> " "ones")"

* LINE 4 "along ...instants" XXXXXXX

 PARAGRAPH 2

* LINE 1 "images,"  ==> "images"

  PARAGRAPH 2

* LINE 3 "The curves allow concluding that" ==> "We conclude that"

 PAGE 4 COLUMN 2

 PARAGRAPH 2

* LINE -1 "to the competitors" ==> "both competitors"

 PARAGRAPH 3

* LINE 1 "This sections demonstrates the WECS" ==> "This sections compares the performance of WECS"

* LINE 2 "methods" XXXXXXX

* LINE 3 "In detail, this application considers" ==> "We consider here

PAGE 5 COLUMN 1

 PARAGRAPH 2

* LINES 3-4 "Figures 9(a)-9(b)" ==> "Figures 6(a)-6(b)"

 PARAGRAPH 3

* LINE 2 "it was" ==> "it is"

 PARAGRAPH 4

* LINE -2 "Figure 7" ==> "Figure 8"

 PAGE 5 COLUMN 2

 PARAGRAPH 2

* LINES 1-3 "After submitting the image series to the methods WECS, ECS, and TAAD, are obtained ... domain" ==> "Through the application WECS, ECS and TAAD to the image series, we obtain matrices R, \tilde{R} and A. They represent the spatially-localized change measures given by WECS, ECS and TAAD, respectively".

PAGE 6 COLUMN 1

 PARAGRAPH 2

* LINE 4 "kappa coefficient with basis" ==> "kappa coefficient based"

PARAGRAPH 3

* LINES 1-2 "the WECS method" ==> "WECS"

 PAGE 6 COLUMN 2

 PARAGRAPH 1

* LINE 5 "the WECS method" ==> "WECS"

 PARAGRAPH 2

* LINE 4 "than ECS and TAAD" ==> "than either ECS or TAAD".

* LINES 4-5 "the TAAD method" ==> "TAAD"

 PARAGRAPH 3

* LINE 1 "the WECS method" ==> "WECS"

PAGE 7 COLUMN 2

 PARAGRAPH 3

* LINE 3-6 "studying energy correlation screening method with for distinct smoothing techniques; deducing change detection rates theoretically for a statistical models" ==> "extending energy correlation screening for distinct smoothing techniques; deducing sharp theoretical change detection rates for appropriate statistical models"

PAGE 3 COLUMN 2

* A sugestão é suprimir "This screening idea has been successfully applied in different problems in areas such as graphical models [31], time series [32] and remote sensing [33]. We show it can also shed light on problems related to change detection in image time series" na página 3. Conseqüentemente, retiram-se as referências [31]-[33] da página 8.

PAGE 1 COLUMN 2

* Trocar "Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task. In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and allows an efficient computational
treatment." por "Change detection process using SAR images is a challenging task, due to the the high-dimensionality of multi-temporal datasets associated with the speckle noise. In this context, the use of Wavelets rises as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms."

PAGE 2 COLUMN 1

* Trocar "The feature screening" por "Feature screening".

PAGE 2 COLUMN 2

* Trocar "that may be involved in wavelet" por "that are involved in wavelet".

Página 1 Coluna 1

Par2 Introduction

* "Many other proposals of methods" => "Many other proposals"

Página 1 Coluna 2

Par2 l5

* ", [18]" ==> XXXXXXX

Par3

* Trocar "Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task. In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and allows an efficient computational
treatment." por "Change detection process using SAR images is a challenging task., due to the the high-dimensionality of multi-temporal datasets associated with the speckle noise. In this context, the use of Wavelets rises as a convenient approach given robustness when dealing with noise data and the available efficient computational algorithms."

Par4

* "Analysis of SAR ... methods, such as [21]-[24]. However," ==>  "SAR ... methods [21]-[24]. On the other hand,"

Página 2 Coluna 1

Par3

* Trocar "The feature screening" por "Feature screening".

* "In order to explain how it works, consider" ==> "Consider"

Página 2 Coluna 2

Par3
* Trocar "that may be involved in wavelet" por "that are involved in wavelet".

Página 3 Coluna 1

Par2

* "Therefore, it is defined the temporal overall variation sequence" ==> "Therefore, the temporal overall variation sequence is defined as"

Página 3 Coluna 2

Par2

* A sugestão é suprimir "This screening idea has been successfully applied in different problems in areas such as graphical models [31], time series [32] and remote sensing [33]. We show it can also shed light on problems related to change detection in image time series" na página 3. Conseqüentemente, retiram-se as referências [31]-[33] da página 8.

Section IV.A

* "A. Experiment design" ==> "A. Experimental design"

Par3

* "proposed method," ==> "proposed method's perfomance,"
* "most appropriated" ==> "most appropriate"

Página 4 Coluna 1

Par1

* "This section applies the WECS, TAAD, and ECS methods to detect changes" ==> "We discuss here the application of  WECS, TAAD, and ECS"

* "plus a noise" ==> "plus noise".

* "first four instants" ==> "first four time instants"

Par2

* "All changes" ==> "Change patterns"

Par3

* "An example of" ==> "A typical example of"

Par4

* "According to the experiment design" ==> "According to the experimental design"

Página 4 Coluna 2

Par1

* "the change and non-changed" ==> "the change and no-change"

Par2

* "considering distint wavelet basis" ==> "considering different wavelet bases"

Par3

* "curves make evident that $J$ equal to 2 or 3 promotes the best performances, with a slight advantage to $J=2$," ==> "curves provides evidence that $J$ equal to 2 or 3 lead to the best performances.  $J=2$ has a slight advantage,

Par4

* "At last, the ROC curves ... 4(c), allow comparing the ... and ECS approaches." ==> "The ROC curves ... 4(c), present the ... and ECS approaches."

* "Firstly, the low-performance ... concluding that the simple swap" ==> "ECS's low performance shows that the simple swap"

* "Reversely, the WECS method provides" ==> "Since WECS provides"

Página 5 Coluna 1

* Fig.4 "ROC curves ... datasey" ==> "ROC curves ... dataset"

Par1

* ", demonstrating then its superiority compared to both competitors" ==> ", its superior performance is clearly established"

Par2

* "are still considered in this analysis" ==> "are employed"

Par4

* "it s possible" ==> "it is possible"

Página 5 Coluna 2

Par2

* "the WECS method" ==> "WECS"

* "central tendency near 12" ==> "central value near 12"

* "than this tendency," ==> "than this baseline,"

* "For sake" ==> "For the sake"

Página 6 Coluna 1

* Fig.7 "Plot ... central tendency around 12" ==> "Plot ... baseline around 12"

* Fig.8 "from distinct instant $m$" ==> ", $m=1, 30, 59$,"

Página 6 Coluna 2

Par2

* "The accuracy of obtained binary maps" ==> "The accuracy of the binary maps"

* "The WECS, ECS and TAAD methods" ==> "WECS, ECS and TAAD"

* "F1-Score and kappa" ==> "the F1-Score and the kappa"

* Fig.9 "assigned to the WECS, ECS and TAAD methods," ==> "associated to WECS, ECS and TAAD,"

Par4

* "Although the performance of WECS equipped" ==> "Although WECS's performance when equipped"

Página 7 Coluna 1

Par1

* "verify that while the TAAD-K1 method assigns" ==> "verify that, while TAAD-K1 assigns"

* " and does not" ==> ", it does not"

* "Regarding the ECS method, the provided result has a noise-corrupt aspect due to the frequent" ==> 
"Regarding ECS, its result has a noise-corrupt character, with the frequent"

Par2

* "the WECS method" ==> "WECS"

Par3

* "corroborate the presented discussions" ==> "corroborate the aforementioned discussion"

* Table 1 "Accuracy ... The Variance of Kappa are multiplied ..." ==> "Accuracy ... Variance of Kappa  multiplied ..."

* Fig10 "Resulting maps" ==> "Resulting change maps"

Página 7 Coluna 2

Par1

* "processing beyond provides more accurate results in comparison to the competitors" ==> 
"processing, beyond providing systematically more accurate results"

Página 8 Coluna 2

* Retirar referências [31]-[33] Zheng at al., Yousuf and Feng, Li et al. .

************************************ IDEIA DE REGRESSAO *****************

Na página 3 coluna 1 o antepenúltimo parágrafo se lê " Define a mapping of {\it relevant} indices ... as ${\mathcal M}_* ...$ where ... in the images".

Aqui, acrescenta-se

"If we apply to Equation (2) the index dicothomy defined by ${\mathcal M}_*$, we have
\begin{equation}
d(m)=\sum_{(k,l)\in\mathcal M_*}\beta_{k,l}D_{k,l}^{(m)}+\varepilon(m), \label{Regressao}
\end{equation}
where $\beta_{k,l}$ are non-null regression coefficients for $(k,l)\in\mathcal M_*$, and $\varepilon(m)$ are stochastic error terms.

The error terms allow for both the apportionment of spurious correlation for indices $notin \mathcal M_*$ as well as for the energies not represented by the wavelet smoothing.

It is a well-known property of discrete wavelet transforms that it statistically decorrelates the original data [19,]-[20]. For instance, this motivates the use of WECS instead of ECS or TAAD, since wavelets will result in sparser representations for $\mathcal M_*$ against either method. Moreover, the sure screening theoretical results for independent data motivates our conjecture that the regression set-up given by Equation (\ref{Regressao}) should have a good performance. A rigorous proof for dependent data sets such as multitemporal series of satellite images is beyond the scope of this manuscript, but the numerical results provide us with some solid evidences."